<div>

---
title: "HARVARD EXTENSION SCHOOL"
subtitle: "EXT CSCI E-106 Model Data Class Group Project Template"
author:
  - "Author: Dinesh Bedathuru"
  - "Author: Brian Calderon"
  - "Author: Jeisson Hernandez"
  - "Author: Hao Fu"
  - "Author: Derek Rush"
  - "Author: Jeremy Tajonera"
  - "Author: Catherine Tully"
date: "`r format(Sys.time(), '%d %B %Y')`"
format:
  pdf:
    toc: true
    number-sections: true
    toc-depth: 2
    keep-tex: true
    geometry: margin=1.3cm
    df-print: kable
tags: [logistic, neuronal networks, titanic]
abstract: |
  In this project, our aim is to classify the probability of a passenger surviving the Titanic crash of 1912. We used a variety of linear and non-linear models to deduce the most accurate model and provide long-term stability in our predictions.
editor_options:
  markdown:
    wrap: 72
---

</div>

```{r echo=FALSE}
#############################################################
#                     Load Data                             #
#############################################################
odata <- read.csv("data/Titanic_Survival_Data.csv")
cat("Size of entire data set:", nrow(odata), "\n")
```

::: pagebreak
:::

Classify whether a passenger on board the maiden voyage of the RMS
Titanic in 1912 survived given their age, sex and class.
Sample-Data-Titanic-Survival.csv to be used in the Final Project

| Variable | Description |
|----|----|
| pclass | **Passenger Class (1 = 1st; 2 = 2nd; 3 = 3rd)** |
| survived | **Survival (0 = No; 1 = Yes)** |
| name | **Name** |
| sex | **Sex** |
| age | **Age** |
| sibsp | **\# of siblings / spouses aboard the Titanic** |
| parch | **\# of parents / children aboard the Titanic** |
| ticket | **Ticket number** |
| fare | **Passenger fare** |
| cabin | **Cabin number** |
| embarked | **Port of Embarkation (C = Cherbourg; Q = Queenstown; S = Southampton)** |
| boat | **Lifeboat ID, if passenger survived** |
| body | **Body number (if passenger did not survive and body was recovered** |
| home.dest | **The intended home destination of the passenger** |

::: pagebreak
:::

# Instructions: 

0\. Join a team with your fellow students with appropriate size (Up to
Nine Students total) If you have not group by the end of the week of
April 11 you may present the project by yourself or I will randomly
assign other stranded student to your group. I will let know the final
groups in April 11.

1\. Load and Review the dataset named “Titanic_Survival_Data.csv” 2.
Create the train data set which contains 70% of the data and use
set.seed (15). The remaining 30% will be your test data set.

3\. Investigate the data and combine the level of categorical variables
if needed and drop variables as needed. For example, you can drop id,
Latitude, Longitude, etc.

4\. Build appropriate model to predict the probability of survival.

5\. Create scatter plots and a correlation matrix for the train data
set. Interpret the possible relationship between the response.

6\. Build the best models by using the appropriate selection method.
Compare the performance of the best logistic linear models.

7\. Make sure that model assumption(s) are checked for the final model.
Apply remedy measures (transformation, etc.) that helps satisfy the
assumptions.

8\. Investigate unequal variances and multicollinearity.

9\. Build an alternative to your model based on one of the following
approaches as applicable to predict the probability of survival:
logistic regression, classification Tree, NN, or SVM. Check the
applicable model assumptions. Explore using a negative binomial
regression and a Poisson regression.

10\. Use the test data set to assess the model performances from above.

11\. Based on the performances on both train and test data sets,
determine your primary (champion) model and the other model which would
be your benchmark model.

12\. Create a model development document that describes the model
following this template, input the name of the authors, Harvard IDs, the
name of the Group, all of your code and calculations, etc..

**Due Date: May 12 2025 1159 pm hours EST Notes No typographical errors,
grammar mistakes, or misspelled words, use English language All tables
need to be numbered and describe their content in the body of the
document All figures/graphs need to be numbered and describe their
content All results must be accurate and clearly explained for a casual
reviewer to fully understand their purpose and impact Submit both the
RMD markdown file and PDF with the sections with appropriate
explanations. A more formal.**

::: pagebreak
:::

# Executive Summary

This section will describe the model usage, your conclusions and any
regulatory and internal requirements. In a real world scneario, this
section is for senior management who do not need to know the details.
They need to know high level (the purpose of the model, limitations of
the model and any issues).

::: pagebreak
:::

# I. Introduction (5 points)

This section needs to introduce the reader to the problem to be
resolved, the purpose, and the scope of the statistical testing applied.
What you are doing with your prediction? What is the purpose of the
model? What methods were trained on the data, how large is the test
sample, and how did you build the model?

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is
displayed).
