<div>

---
title: "HARVARD EXTENSION SCHOOL"
format: pdf
subtitle: "EXT CSCI E-106 Model Data Class Group Project Template"
author:
- Author Dinesh Bedathuru
- Author Brian Calderon
- Author Jeisson Hernandez
- Author Hao Fu
- Author Derek Rush
- Author Jeremy Tajonera
- Author Catherine Tully

tags: [logistic, neuronal networks, titanic]
abstract: |
   In this project our aim is to classisfy the probability of a passenger to survive the Titanic crash of 1912. We used a variety of linear and non-linear to deduce the most accurate model and provide the best long term stability in our predictions. 
   
date: "`r format(Sys.time(), '%d %B %Y')`"
geometry: margin=1.3cm
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

</div>

## Quarto

## Running Code

When you click the **Render** button a document will be generated that
includes both content and the output of embedded code. You can embed
code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is
displayed).
